Verification of Stencil Computations with Spatial Specifications

### 2016-11-11 paper.tex

Regions comprise \emph{region constants} which are 
combined via operators \term{+} and \term{*}.

@@

@@

\begin{definition}
  An \emph{index scheme} is a vector of size $n$ (an $n$-vector) with
  values drawn from $\domainVal = \mathbb{Z} \cup \{\infty\}$.  Values
  in $\mathbb{Z}$ represent constant offsets of neighbourhood indices
  (expressions of the form $i \pm c$ where $i$ is an induction
  variable and $c$ is a constant integer, Def.~\ref{def:neighbour},
  p.~\pageref{def:neighbour}). $\infty$ represents any indexing
  behaviour--- it is a wildcard.  Throughout $u, v, w$ range over
  index schemes and $u_i$ denotes the $i^{th}$ element of $u$. We
  write $(\domainVal)^n$ for the set of $n$-dimensional schemes.
\end{definition}

@@

\todo{Consider what to do regarding stencil model
- can give previous model, or can try to base on the work cited
on Wikipeadia that we discussed.}

@@

The intuition behind $\otimes$ is that it takes
all possible pairs of indexing schemes,
and treats each pair as lower and upper bounds of an 
$n$-dimensional rectangle. From these lower and upper bounds

 treating them as the
lower and upper bounds of an $n$-dimensional rectangle from which the
remaining vertices of the rectangle are generated (like a bounding
box) \eg{}:
%%
\begin{align*}
\{\vtwohs{1}{2}, \vtwohs{5}{6}\,\} \otimes \{\vtwohs{3}{4}\} =
\{\vtwohs{1}{2}, \vtwohs{3}{2}, \vtwohs{1}{4}, \vtwohs{3}{4}, 
   \vtwohs{3}{6}, \vtwohs{5}{4}, \vtwohs{5}{6}\,\}
\end{align*}
%%
The tensor takes a Cartesian product of the two models,
giving all pairs of schemes which are combined using the
\emph{pairwise permutation} written $\bowtie$ (defined later).

@@

>>>>

Regions comprise \emph{region constants} which are 
combined via region operators \term{+} and \term{*}.

@@

\paragraph{Combining regions}

The region operators \term{+} and \term{*} provide ways to combine
regions in a disjunctive and conjuctive fashion respectively.

The conjuction of two regions $r \term{*} s$ means that any indices
in the specified code must be consistent with both $r$ and $s$ at
the same time. Dually, there must be an index which is consitent
with every component of region $r$ and

@@

\begin{definition}
  An \emph{index scheme} is a vector of size $n$ (an $n$-vector) with
  values drawn from $\domainVal = \mathbb{Z} \cup \{\infty\}$.  Values
  in $\mathbb{Z}$ represent the offsets of neighbourhood indices
  (\ie{}, the constant integer $c$ in an indexing expression of the
  form $i \pm c$, Def.~\ref{def:neighbour},
  p.~\pageref{def:neighbour}) and $\infty$ represents any constant
  inddex. Throughout $u, v, w$ range over
  index schemes and $u_i$ denotes the $i^{th}$ element of $u$. We
  write $(\domainVal)^n$ for the set of $n$-dimensional schemes.
\end{definition}

@@

\begin{definition}[Quotienting schemes]
A scheme $u$ can be quotiented by a set of dimension identifies
by the operator $/_{n,I}$ where $I$ is a set of dimension
identifiers drawn from $\mathbb{N}_{\leq n}$ where 
\begin{equation*}
u/_{n,I} = [ ]
\end{equation*}
\end{definition}

@@

The intuition behind $\otimes$ is that it takes all possible pairs of
indexing schemes, and treats each pair as lower and upper bounds of an
$n$-dimensional rectangle. From these lower and upper bounds, 
the remaining vertices of the $n$-dimensional rectangle are
generated (like a bounding box) \eg{}:
%%
\begin{align*}
\{\vtwohs{1}{2}, \vtwohs{5}{6}\,\} \otimes \{\vtwohs{3}{4}\} =
\{\vtwohs{1}{2}, \vtwohs{3}{2}, \vtwohs{1}{4}, \vtwohs{3}{4}, 
   \vtwohs{3}{6}, \vtwohs{5}{4}, \vtwohs{5}{6}\,\}
\end{align*}
%%
The tensor takes a Cartesian product of the two models,
giving all pairs of schemes which are combined using the
\emph{pairwise permutation} written $\bowtie$ (defined later).
Pairwise permutation generates all possible interleavings of two
vectors. Some care is then taken over
how this interacts with $\infty$: pairwise permutation
corresponds to non-deterministic choice, in each dimension $i$,
between two values $u_i$ and $v_i$, but if either
is $\infty$ then the resulting pairwise permutation is
filtered out if the dimension $i$ has \emph{any other scheme
in the model where $u_i \neq \infty$}, \eg{}:
\begin{align*}
\{\vtwoh{1}{\infty}\} \otimes \{\vtwoh{3}{4}\} =
\{\vtwoh{1}{4}, \vtwoh{3}{4}\}
\end{align*}
%
where $\vtwoh{1}{\infty}$ and $\vtwoh{3}{\infty}$ have been filtered.

The notion of which dimensions $i$ have \emph{any scheme in the model
where $u_i \neq \infty$} is represented by the \emph{constrained
dimensions} intermediate function:
\begin{definition}%[Constrained dimensions]
For an $n$-dimensional model $M$, its \emph{constrained dimensions}
is the set of dimension identifiers for which at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension. That is:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}
\begin{definition}The tensor $\otimes_n$ of $n$-dimensional
models is defined:
\begin{alignat*}{2}
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{5.2cm}{$(u, v) \in M \times N $ \\[0.1em]
                 $\;\; \wedge \;\, j \in (\effdims{M}{n} \cup
                   \effdims{N}{n})$\\[0.1em]
                  $\;\; \wedge \;\, i \in \{1, \ldots, 2^n\} \, \wedge
                  \, x = (u \bowtie v)_i$ \\[0.1em]
                   $\;\; \wedge \;\, x_j \neq \infty$
                  } \Bigg\}
\end{alignat*}
The first guard takes the Cartesian product of
the models and quantifies over all pairings $(u, v)$.
The second guard binds $j$ which ranges over the constrained
dimensions for $M$ unioned with the constrained dimensions for $N$.
That is, every $j$ corresponds to a dimension in which
either $M$ or $N$ has a scheme $u$ where $u_j \neq \infty$,
and thus is constrained.
The \emph{pairwise permutation} $u \bowtie v$ 
builds a $2^n \times n$ matrix 
of all possible $n$-vectors generated by
non-deterministically picking for each $j^{th}$ entry either 
$u_j$ or $v_j$ \eg{}
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. The pairwise permutation
matrix is thus defined in terms of the logical
matrix $b$ where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$ as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
The permuted schemes $x = (u \bowtie v)_i$ added to the model 
$M \otimes N$ are only those which have do not have $\infty$ in 
the constrained dimensions given by each model ($x_j \neq
\infty$). This filters out all permutations which have
$\infty$ in a constrained dimension position. 
\end{definition}

@@

%\newcommand{\cons}[3]{\mathbf{C}_{#3}(#1, #2)}
\newcommand{\cons}[3]{#2 \;\, \textbf{consistent} \,\, #1}
\begin{align*}
\dfrac{\forall u \in C, \exists v \in S .  \, u \preceq v}
      {\cons{\mathsf{up}(S)}{C}{n}}
\qquad
\dfrac{\forall v \in S, \exists u \in C . \, u \preceq v}
      {\cons{\mathsf{low}(S)}{C}{n}} \\[0.75em]
\dfrac{\cons{\mathsf{down}(S)}{C}{n} \qquad \cons{\mathsf{up}(S)}{C}{n}}
  {\cons{\mathsf{exact}(S)}{C}{n}} \\[0.75em]
\dfrac{\cons{\mathsf{down}(S_1)}{C}{n} \qquad \cons{\mathsf{up}(S_2)}{C}{n}}
  {\cons{\mathsf{both}(S_1, S_2)}{C}{n}}
\end{align*}
